How is experience transformed into memory?

### 2018-09-05 paper/main.tex
backing up edits to supplemental materials [WIP]
renamed supplemental materials supporting information

\newcommand{\topics}{S1}
\newcommand{\corrmats}{S2}
\newcommand{\matchmats}{S3}
\newcommand{\featureimportance}{S4}
\newcommand{\listlearning}{S5}
\newcommand{\topicopt}{S6}

@@

Because our analysis framework projects the dynamic video content and participants' recalls onto a shared topic space, and because the dimensions of that space are known (i.e., each topic dimension is a set of weights over words in the vocabulary; Fig.~\topics), we can examine the topic trajectories to understand which specific content was remembered well (or poorly).  For each video event, we can ask: what was the average correlation (across participants) between the video event's topic vector and the closest matching recall event topic vectors from each participant? This yields a single correlation coefficient for each video event, describing how closely participants' recalls of the event tended to reliably capture its content (Fig.~\ref{fig:topics}A).  (We also examined how different comparisons between each video event's topic vector and the corresponding recall event topic vectors related to hand-annotated characterizations of memory performance; see \textit{Supplemental Materials}).  Given this summary of which events were recalled reliably (or not), we next asked whether the better-remembered or worse-remembered events tended to reflect particular topics.  We computed a weighted average of the topic vectors for each video event, where the weights reflected how reliably each event was recalled.  To visualize the result, we created a ``wordle'' image~\citep{MuelEtal18} where words weighted more heavily by better-remembered topics appear in a larger font (Fig.~\ref{fig:topics}B, green box).  Events that reflected topics weighting heavily on characters like ``Sherlock'' and ``John'' (i.e., the main characters) and locations like ``221b Baker Street'' (i.e., a major recurring location; the address of the flat that Sherlock and John share) were best remembered.  An analogous analysis revealed which themes were poorly remembered; here in computing the weighted average over events' topic vectors we weighted each event in \textit{inverse} proportion to how well it was remembered (Fig.~\ref{fig:topics}B, red box).  This revealed that events with relatively minor characters such as ``Mike,'' ``Jeffrey,'' and ``Molly,'' as well as less-integral plot locations (e.g., ``hospital'' and ``office'') were least well-remembered.  This suggests that what is retained in memory are the major plot elements (i.e., the overall ``gist'' of what happened), whereas the more minor details are prone to pruning.

@@

We created similar topic proportion matrices using hand-annotated transcripts of each participant's recall of the video~\citep[annotated by ][]{ChenEtal17}.  We tokenized the transcript into a list of sentences, and then re-organized the list into overlapping sliding windows spanning 10 sentences each; we in turn transformed each window's sentences into a word count vector (using the same vocabulary as for the video model).  We then used the topic model already trained on the video scenes to compute the most probable topic proportions for each sliding window.  This yielded a number-of-sentences (range: 68--294) by number-of-topics (100) topic proportions matrix, for each participant.  These reflected the dynamic content of each participant's recalls.  Note: for details on how we selected the video and recall window lengths and number of topics, see \textit{Supplemental Materials} and Figure~\topicopt.

>>>>

\newcommand{\topicopt}{S1}
\newcommand{\topics}{S2}
\newcommand{\featureimportance}{S3}
\newcommand{\listlearning}{S4}
\newcommand{\corrmats}{S5}
\newcommand{\matchmats}{S6}

@@

Because our analysis framework projects the dynamic video content and participants' recalls onto a shared topic space, and because the dimensions of that space are known (i.e., each topic dimension is a set of weights over words in the vocabulary; Fig.~\topics), we can examine the topic trajectories to understand which specific content was remembered well (or poorly).  For each video event, we can ask: what was the average correlation (across participants) between the video event's topic vector and the closest matching recall event topic vectors from each participant? This yields a single correlation coefficient for each video event, describing how closely participants' recalls of the event tended to reliably capture its content (Fig.~\ref{fig:topics}A).  (We also examined how different comparisons between each video event's topic vector and the corresponding recall event topic vectors related to hand-annotated characterizations of memory performance; see \textit{Supporting information}).  Given this summary of which events were recalled reliably (or not), we next asked whether the better-remembered or worse-remembered events tended to reflect particular topics.  We computed a weighted average of the topic vectors for each video event, where the weights reflected how reliably each event was recalled.  To visualize the result, we created a ``wordle'' image~\citep{MuelEtal18} where words weighted more heavily by better-remembered topics appear in a larger font (Fig.~\ref{fig:topics}B, green box).  Events that reflected topics weighting heavily on characters like ``Sherlock'' and ``John'' (i.e., the main characters) and locations like ``221b Baker Street'' (i.e., a major recurring location; the address of the flat that Sherlock and John share) were best remembered.  An analogous analysis revealed which themes were poorly remembered; here in computing the weighted average over events' topic vectors we weighted each event in \textit{inverse} proportion to how well it was remembered (Fig.~\ref{fig:topics}B, red box).  This revealed that events with relatively minor characters such as ``Mike,'' ``Jeffrey,'' and ``Molly,'' as well as less-integral plot locations (e.g., ``hospital'' and ``office'') were least well-remembered.  This suggests that what is retained in memory are the major plot elements (i.e., the overall ``gist'' of what happened), whereas the more minor details are prone to pruning.

@@

We created similar topic proportion matrices using hand-annotated transcripts of each participant's recall of the video~\citep[annotated by ][]{ChenEtal17}.  We tokenized the transcript into a list of sentences, and then re-organized the list into overlapping sliding windows spanning 10 sentences each; we in turn transformed each window's sentences into a word count vector (using the same vocabulary as for the video model).  We then used the topic model already trained on the video scenes to compute the most probable topic proportions for each sliding window.  This yielded a number-of-sentences (range: 68--294) by number-of-topics (100) topic proportions matrix, for each participant.  These reflected the dynamic content of each participant's recalls.  Note: for details on how we selected the video and recall window lengths and number of topics, see \textit{Supporting information} and Figure~\topicopt.
